Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2023-05-08 doc/manuscript/ESSDmanuscript.Rmd
incorporate updated results
 #42

n_ind_records_EFDB_variables= C_variables[length(C_variables[,"variable"]), "n independent records in ForC"] #n independent records relevant to EFDB
n_ind_records_agb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag"]
n_ind_records_agb_components <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_foliage"]
n_ind_records_biomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass"]
n_ind_records_rootbiomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_coarse"]
n_ind_records_deadwood <- C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_down"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_standing"]
n_ind_records_OL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "litter"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "O.horizon"]
n_ind_records_SOC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "SOM / SOC"] 
n_ind_records_deltaagb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.agb"] 
n_ind_records_deltaOL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.O.horizon"] 
flux_variables_with_records <- c("NPP_woody", "ANPP_woody", "ANPP_woody_ag", "BNPP_root_fine","BNPP_root.turnover_fine","BNPP_root_coarse" )
n_ind_records_flux <- sum(C_variables$`n independent records in ForC`[C_variables$`variable` %in% flux_variables_with_records])

@@

ForC contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables relevant to EFDB (Table `r table_variables`, Fig. `r fig_variable_mapping`).
The records were very unevenly distributed across variables. 
The variable with most records was aboveground biomass, representing `r round(n_ind_records_agb/n_ind_records_EFDB_variables,2)*100`% of all independent records relevant to EFDB, and aboveground biomass components (woody biomass or foliage) representing an additional `r round(n_ind_records_agb_components/n_ind_records_EFDB_variables,2)*100`%.
A total of `r round(n_ind_records_rootbiomass/n_ind_records_EFDB_variables,2)*100`% of relevant records were for root biomass (including fine and coarse root components), while `r round(n_ind_records_biomass/n_ind_records_EFDB_variables,2)*100`% described total biomass.
The non-living pools were less represented, with `r round(n_ind_records_deadwood/n_ind_records_EFDB_variables,2)*100`% of relevant were for dead wood (including standing and fallen components), `r round(n_ind_records_OL/n_ind_records_EFDB_variables*100,1)`% for litter, and `r round(n_ind_records_SOC/n_ind_records_EFDB_variables*100,1)`% for soil carbon.

Increment and flux variables were poorly represented (Table `r table_variables`).
The increment variable with most records was the aboveground biomass increment, representing `r round(n_ind_records_deltaagb/n_ind_records_EFDB_variables*100,1)`% of all independent records relevant to EFDB.
The only other relevant increment variable with any records was the O horizon (litter) increment, with just `r n_ind_records_deltaOL` records.
Relevant flux variable records (n=`r n_ind_records_flux`) were limited to the biomass pools (aboveground, belowground, or total) and together constituted `r round(n_ind_records_flux/n_ind_records_EFDB_variables,2)*100`% of ForC's independent records relevant to EFDB.

@@

<!-- this paragraph will need to be revisited after we solve issues 42 and 43-->
The records transferred from ForC to EFDB were broadly distributed across Earth's forests (Fig. `r fig_map`).
However, the density of these records was very unevenly distributed across continents, biomes, and forest types and was not proportional to the numbers of relevant records in ForC (Fig. `r fig_histograms`).
Rather, the largest number of records came from North America, followed by Asia and South America (Fig. `r fig_histograms`c), with the most represented FAO ecozones being boreal coniferous forest, temperate continental forest, and temperate mountain systems, followed by tropical rain forests and moist deciduous forests (Fig. `r fig_histograms`b).
In terms of dominant vegetation, by far the most records came from needleleaf evergreen forests, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`b). 
The largest records came from mature forests (>100 years), followed by young and intermediate-aged stands (Fig. `r fig_histograms`d).

>>>>

n_ind_records_EFDB_variables= C_variables[length(C_variables[,"variable"]), "n independent records in ForC"] #n independent records relevant to EFDB
n_ind_records_agb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag"]
n_ind_records_agb_components <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_foliage"]
n_ind_records_biomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass"]
n_ind_records_rootbiomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_coarse"]
n_ind_records_deadwood <- C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_down"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_standing"]
n_ind_records_OL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "litter"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "O.horizon"]
n_ind_records_TotEcosystemC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "total.ecosystem_2"] 
n_ind_records_SOC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "SOM / SOC"] 
n_ind_records_deltaagb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.agb"] 
n_ind_records_deltaOL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.O.horizon"] 
flux_variables_with_records <- c("NPP_woody", "ANPP_woody", "ANPP_woody_stem", "ANPP_woody_branch", "woody.mortality_ag", "BNPP_root", "BNPP_root_fine","BNPP_root.turnover_fine","BNPP_root_coarse" , "ANPP_litterfall", "NPP_litter", "R_het_litter", "R_het_soil")
n_ind_records_flux <- sum(C_variables$`n independent records in ForC`[C_variables$`variable` %in% flux_variables_with_records])

@@

ForC contained records for `r n_relevant_variables_with_data` of the `r n_relevant_variables` variables relevant to EFDB (Table `r table_variables`, Fig. `r fig_variable_mapping`).
The records were very unevenly distributed across variables. 
The variable with most records was aboveground biomass, representing `r round(n_ind_records_agb/n_ind_records_EFDB_variables,2)*100`% of all independent records relevant to EFDB, and aboveground biomass components (woody biomass or foliage) representing an additional `r round(n_ind_records_agb_components/n_ind_records_EFDB_variables,2)*100`%.
A total of `r round(n_ind_records_rootbiomass/n_ind_records_EFDB_variables,2)*100`% of relevant records were for root biomass (including fine and coarse root components), while `r round(n_ind_records_biomass/n_ind_records_EFDB_variables,2)*100`% described total biomass.
The non-living pools were less represented, with `r round(n_ind_records_deadwood/n_ind_records_EFDB_variables,2)*100`% of relevant were for dead wood (including standing and fallen components), `r round(n_ind_records_OL/n_ind_records_EFDB_variables*100,1)`% for litter, `r round(n_ind_records_TotEcosystemC/n_ind_records_EFDB_variables*100,1)`% for total ecosystem C excluding soils, and `r round(n_ind_records_SOC/n_ind_records_EFDB_variables*100,1)`% for soil carbon.

Increment and flux variables were poorly represented (Table `r table_variables`).
The increment variable with most records was the aboveground biomass increment, representing `r round(n_ind_records_deltaagb/n_ind_records_EFDB_variables*100,1)`% of all independent records relevant to EFDB.
The only other relevant increment variable with any records was the O horizon (litter) increment, with just `r n_ind_records_deltaOL` records.
Relevant flux variable records (n=`r n_ind_records_flux`) together constituted `r round(n_ind_records_flux/n_ind_records_EFDB_variables,2)*100`% of ForC's independent records relevant to EFDB.

@@

The records transferred from ForC to EFDB were broadly distributed across Earth's forests (Fig. `r fig_map`).
However, the density of these records was very unevenly distributed across continents, biomes, and forest types and was not proportional to the numbers of relevant records in ForC (Fig. `r fig_histograms`).
Rather, the largest number of records came from North America, followed by Asia, South America, and Africa (Fig. `r fig_histograms`c), with the most represented FAO ecozones being boreal coniferous forest, temperate continental forest, and temperate mountain systems, followed by tropical rain forests and moist deciduous forests (Fig. `r fig_histograms`b).
In terms of dominant vegetation, by far the most records came from needleleaf evergreen forests, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`b). 
The largest records came from mature forests (>100 years), followed by young and intermediate-aged stands (Fig. `r fig_histograms`d).
